---
title: "seurat-analysis"
author: "Elinor Wing"
date: "20/05/2022"
output: html_document
---

EW-2

Start with 13572 cells
-11848 after removing high mit/RNA
-3 CD14+
-5 LYZ+
-11840 after removing contaminating cells


```{r}
library(dplyr)
library(Seurat)
library(patchwork)

# Load the dataset
EW.2 <- Read10X(data.dir = "./sample2_filtered_feature_bc_matrix/")
# Initialize the Seurat object with the raw (non-normalized data).
EW.2 <- CreateSeuratObject(counts = EW.2, project = "synovial-b-cells", min.cells = 3, min.features = 200)
EW.2
```


## QC and selecting cells for further analysis

```{r}
EW.2[["percent.mt"]] <- PercentageFeatureSet(EW.2, pattern = "^MT-")
```

```{r}
EW.2[["percent.rb"]] <- PercentageFeatureSet(EW.2, pattern = "^RP[SL]")
```



```{r}
# Visualize QC metrics as a violin plot
VlnPlot(EW.2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4)
```


```{r}
plot1 <- FeatureScatter(EW.2, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(EW.2, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```

```{r}
EW.2 <- subset(EW.2, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)

EW.2
```



```{r}
#remove contaminating cells
EW.2 <- subset(EW.2, subset = CD3E >= 1, invert =TRUE) #T cells
EW.2 <- subset(EW.2, subset = CD14 >= 1, invert =TRUE) #monocytes
EW.2 <- subset(EW.2, subset = GNLY >= 1, invert =TRUE) #CTLs
EW.2 <- subset(EW.2, subset = FCER1A >= 1, invert =TRUE) #mast cells
EW.2 <- subset(EW.2, subset = GCGR3A >= 1, invert =TRUE) #CD16
EW.2 <- subset(EW.2, subset = LYZ >= 1, invert =TRUE) #mono and myeloid DC
EW.2 <- subset(EW.2, subset = PPBP >= 1, invert =TRUE) #platelet

EW.2
```



## Normalizing the data

```{r}
EW.2 <- NormalizeData(EW.2)
```


## Identification of highly variable features

```{r}
EW.2 <- FindVariableFeatures(EW.2, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(EW.2), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(EW.2)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
plot2
```


## Scaling the data

https://www.genenames.org/data/genegroup/#!/group/348

```{r}
#lists of BCR genes
#would not allow me to combine. adding any more to regress.genes1 created an error
regress.genes1 <- list(c("IGHE", "IGHA1", "IGHA2", "IGHM", "IGHG1", "IGHG2", "IGHG3", "IGHG4", "IGHD", "IGKV1-5", "IGKV1-6", "IGKV1-8", "IGKV1-9", "IGKV1-12", "IGKV1-13", "IGKV1-16", "IGKV1-17", "IGKV1-22", "IGKV1-27", "IGKV1-32", "IGKV1-33", "IGKV1-35", "IGKV1-37", "IGKV1-39", "IGKV1D-8", "IGKV1D-12", "IGKV1D-13", "IGKV1D-16", "IGKV1D-17", "IGKV1D-22", "IGKV1D-27", "IGKV1D-32", "IGKV1D-33", "IGKV1D-35", "IGKV1D-37", "IGKV1D-39", "IGKV1D-42", "IGKV1D-43", "IGKV2-4", "IGKV2-10", "IGKV2-14", "IGKV2-18", "IGKV2-19", "IGKV2-23", "IGKV2-24", "IGKV2-26", "IGKV2-28", "IGKV2-29", "IGKV2-30", "IGKV2-36", "IGKV2-38", "IGKV2-40", "IGKV2D-10", "IGKV2D-14", "IGKV2D-18", "IGKV2D-19", "IGKV2D-23", "IGKV2D-24", "IGKV2D-26", "IGKV2D-28", "IGKV2D-29", "IGKV2D-30", "IGKV2D-36", "IGKV2D-38", "IGKV2D-40", "IGKV3-7", "IGKV3-11", "IGKV3D-7", "IGKV3-15", "IGKV3D-11", "IGKV4-1", "IGKV3-20", "IGKV3D-15", "IGKV5-2", "IGKV3-25", "IGKV3D-20", "IGKV6-21", "IGKV3-31", "IGKV3D-25", "IGKV6D-21", "IGKV6D-21", "IGKV3-34", "IGKV3D-31", "IGKV6D-41", "IGKV3D-34", "IGKV7-3", "IGKJ1", "IGKJ2", "IGKJ3", "IGKJ4", "IGKJ5", "IGKC", "IGLV1-36", "IGLV1-40", "IGLV1-41", "IGLV1-44", "IGLV1-47", "IGLV1-50", "IGLV1-51", "IGLV1-62", "IGLV2-5", "IGLV2-8", "IGLV2-11", "IGLV2-14", "IGLV2-18", "IGLV2-23", "IGLV2-28", "IGLV2-33", "IGLV2-34", "IGLV3-1", "IGLV3-2", "IGLV3-4", "IGLV3-6", "IGLV3-7", "IGLV3-9", "IGLV3-10", "IGLV3-12", "IGLV3-13", "IGLV3-15", "IGLV3-16", "IGLV3-17", "IGLV3-19", "IGLV3-21", "IGLV3-22", "IGLV3-24", "IGLV3-25", "IGLV3-26", "IGLV3-27", "IGLV3-29", "IGLV3-30", "IGLV3-31", "IGLV3-32", "IGLV4-3", "IGLV4-60", "IGLV4-69", "IGLV5-37", "	IGLV5-39", "	IGLV5-45", "IGLV5-48", "IGLV5-52", "IGLV6-57", "IGLV7-35", "IGLV7-43", "IGLV7-46", "IGLV8-61", "IGLV9-49", "IGLV10-54", "IGLV10-67", "IGLV11-55", "IGLVI-20", "IGLVI-38", "IGLVI-42", "IGLVI-56", "IGLVI-63", "IGLVI-68", "IGLVI-70", "IGLVIV-53", "IGLVIV-59", "IGLVIV-64", "IGLVIV-65", "IGLVV-58", "IGLVV-66", "	IGLVVI-22-1", "IGLVVI-25-1", "IGLVVII-41-1", "IGLJ1", "IGLJ2", "IGLJ3", "IGLJ4", "IGLJ5", "IGLJ6", "IGLJ7", "IGLC1", "IGLC2", "IGLC3", "IGLC4", "IGLC5", "IGLC6", "IGLC7", "IGHEP1", "IGHGP", "IGHJ6", "IGHJ5", "IGHJ4", "IGHJ3P", "IGHJ3", "IGHJ2P", "IGHJ2", "IGHJ1P", "IGHJ1", "IGHD7-27", "IGHD6-25", "IGHD6-19", "IGHD6-13", "IGHD6-6", "IGHD5-24", "IGHD5-18", "IGHD5-12", "IGHD5-5", "IGHD4-23", "IGHD4-17", "IGHD4-11", "IGHD4-4", "IGHD3-22", "IGHD3-16", "IGHD3-10", "IGHD3-9", "IGHD3-3", "IGHD2-21", "IGHD2-15", "IGHD2-8", "IGHD2-2", "IGHD1-26", "IGHD1-20", "IGHD1-14", "IGHD1-7", "IGHD1-1", "IGHVIV-44-1", "IGHVIII-82", "IGHVIII-76-1", "IGHVIII-67-4", "IGHVIII-67-3", "IGHVIII-67-2", "IGHV8-51-1", "	IGHVIII-47-1", "	IGHVIII-44", "IGHVIII-38-1", "	IGHVIII-26-1", "IGHVIII-25-1", "IGHVIII-22-2", "IGHVIII-16-1", "IGHVIII-13-1", "IGHVIII-11-1", "IGHVIII-5-2", "IGHVIII-5-1", "IGHVIII-5-2", "IGHVIII-11-1", "IGHVIII-5-2", "IGHVIII-5-1", "IGHVIII-2-1", "IGHVII-78-1", "IGHVII-74-1", "IGHVII-67-1", "IGHVII-65-1", "IGHVII-62-1", "IGHVII-60-1", "IGHVII-53-1", "IGHVII-51-2", "IGHVII-49-1", "IGHVII-46-1", "IGHVII-44-2", "IGHVII-43-1", "IGHVII-40-1", "IGHVII-33-1", "IGHVII-31-1", "IGHVII-30-21", "IGHVII-30-1", "IGHVII-28-1", "IGHVII-26-2", "IGHVII-22-1", "IGHVII-20-1", "IGHVII-15-1", "IGHVII-1-1", "IGHV7-81", "IGHV7-34-1", "IGHV7-56", "IGHV7-27", "IGHV7-40", "IGHV7-4-1", "IGHV6-1", "IGHV5-10-1", "IGHV5-78", "IGHV5-51", "IGHV4-38-2", "IGHV4-80", "IGHV4-61", "IGHV4-59", "IGHV4-55", "IGHV4-39", "IGHV4-34", "IGHV4-31", "IGHV4-30-4", "IGHV4-30-2", "IGHV4-30-1", "IGHV4-28", "IGHV4-4", "IGHV3-69-1", "IGHV3-38-3", "IGHV3-79", "IGHV3-76", "IGHV3-75", "IGHV3-74", "IGHV3-73", "IGHV3-72", "IGHV3-71", "IGHV3-66", "IGHV3-65", "IGHV3-64", "IGHV3-63", "IGHV3-62", "IGHV3-60", "IGHV3-57", "IGHV3-54", "IGHV3-53", "IGHV3-52", "IGHV3-50", "IGHV3-49", "IGHV3-48", "IGHV3-47", "IGHV3-43", "IGHV3-42", "IGHV3-41", "IGHV3-38", "IGHV3-37", "IGHV3-36", "IGHV3-35", "IGHV3-33-2", "IGHV3-33", "IGHV3-32", "IGHV3-30-5", "IGHV3-30-3", "IGHV3-30-2", "IGHV3-30", "IGHV3-29", "IGHV3-25", "IGHV3-23", "IGHV3-22", "IGHV3-21", "IGHV3-20", "IGHV3-19", "IGHV3-16", "IGHV3-15")) 

regress.genes2 <- list(c("IGHV3-13", "IGHV3-11", "IGHV3-9", "IGHV3-7", "IGHV3-6", "IGHV2-70D", "IGHV2-70", "IGHV2-26", "IGHV2-10", "IGHV2-5", "IGHV1-69-2", "IGHV1-38-4", "IGHV1-69D", "IGHV1-69", "IGHV1-68", "IGHV1-67", "IGHV1-58", "IGHV1-46", "IGHV1-45", "IGHV1-24", "IGHV1-18", "IGHV1-17", "IGHV1-14", "IGHV1-8", "IGHV1-12", "IGHV1-3", "IGHV1-2", "IGHV3-64D"))
```


```{r}
#for loop to interate through the variable genes list and remove any of the BCR related genes in the lists above
for ( i in EW.2@assays[["RNA"]]@var.features){
  if (i%in%(regress.genes1[[1]])){EW.2@assays[["RNA"]]@var.features <- EW.2@assays[["RNA"]]@var.features[which(EW.2@assays[["RNA"]]@var.features != i)]}
}
```


```{r}
for ( i in EW.2@assays[["RNA"]]@var.features){
  if (i%in%(regress.genes2[[1]])){EW.2@assays[["RNA"]]@var.features <- EW.2@assays[["RNA"]]@var.features[which(EW.2@assays[["RNA"]]@var.features != i)]}
}
```

```{r}
EW.2 <- ScaleData(EW.2, features = rownames(EW.2))
```


## 7. Identify cell-cycle scores

```{r}
# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
```

```{r}
EW.2 <- CellCycleScoring(EW.2, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

# view cell cycle scores and phase assignments
head(EW.2[[]])
```

## Perform linear dimensional reduction

```{r}
EW.2 <- RunPCA(EW.2, features = VariableFeatures(object = EW.2))
```

```{r}
VizDimLoadings(EW.2, dims = 1:2, reduction = "pca")
```

```{r}
DimHeatmap(EW.2, dims = 1:15, cells = 500, balanced = TRUE)
```

## Determine the ‘dimensionality’ of the dataset

```{r}
EW.2 <- JackStraw(EW.2, num.replicate = 100)
EW.2 <- ScoreJackStraw(EW.2, dims = 1:20)
```

```{r}
JackStrawPlot(EW.2, dims = 1:20)
```


```{r}
ElbowPlot(EW.2)
```


## Cluster

```{r}
EW.2 <- FindNeighbors(EW.2, dims = 1:15)

#Identify clusters of cells by a shared nearest neighbor (SNN) modularity optimization based clustering algorithm.
EW.2 <- FindClusters(EW.2, resolution = 0.7)
```


## Run non-linear dimensional reduction (UMAP/tSNE)


```{r}
EW.2 <- RunUMAP(EW.2, dims = 1:15)
```

```{r}
DimPlot(EW.2, reduction = "umap")
```


```{r}
#number of reads
VlnPlot(EW.2,features="nCount_RNA")
```

```{r}
#number of reads
FeaturePlot(EW.2, features = "nCount_RNA", label.size = 4, repel = T, label = T) & theme(plot.title = element_text(size=10))
```

```{r}
#number of genes
VlnPlot(EW.2,features="nFeature_RNA")
```

```{r}
#number of genes
FeaturePlot(EW.2, features = "nFeature_RNA", label.size = 4, repel = T, label = T) & theme(plot.title = element_text(size=10))
```

```{r}
#percent mitochondrial genes
VlnPlot(EW.2,features="percent.mt", pt.size = 0.005)
```

```{r}
#percent mitochondrial genes
FeaturePlot(EW.2, features = "percent.mt", label.size = 4, repel = T, label = T) & theme(plot.title = element_text(size=10))
```

```{r}
#percent ribosomal protein genes
VlnPlot(EW.2,features="percent.rb")
```

```{r}
#percent ribosomal protein genes
FeaturePlot(EW.2, features = "JCHAIN", label.size = 4, repel = T, label = T) & theme(plot.title = element_text(size=10))
```

```{r}
#cell-cycle scores
VlnPlot(EW.2, features = c("S.Score","G2M.Score"))
```

```{r}
#cell-cycle scores
FeaturePlot(EW.2 ,features = c("S.Score","G2M.Score"),label.size = 4,repel = T,label = T) & 
  theme(plot.title = element_text(size=10))
```


```{r}
cluster0.markers <- FindMarkers(EW.2, ident.1 = 0, min.pct = 0.25)
write.csv(cluster0.markers,"./EW2-out/cluster0.markers.csv", row.names = TRUE)
head(cluster0.markers, n = 6)
```

```{r}
cluster1.markers <- FindMarkers(EW.2, ident.1 = 1, min.pct = 0.25)
write.csv(cluster1.markers,"./EW2-out/cluster1.markers.csv", row.names = TRUE)
head(cluster1.markers, n = 6)
```

```{r}
cluster2.markers <- FindMarkers(EW.2, ident.1 = 2, min.pct = 0.25)
write.csv(cluster2.markers,"./EW2-out/cluster2.markers.csv", row.names = TRUE)
head(cluster2.markers, n = 6)
```

```{r}
cluster3.markers <- FindMarkers(EW.2, ident.1 = 3, min.pct = 0.25)
write.csv(cluster3.markers,"./EW2-out/cluster3.markers.csv", row.names = TRUE)
head(cluster3.markers, n = 6)
```

```{r}
cluster4.markers <- FindMarkers(EW.2, ident.1 = 4, min.pct = 0.25)
write.csv(cluster4.markers,"./EW2-out/cluster4.markers.csv", row.names = TRUE)
head(cluster4.markers, n = 6)
```

```{r}
cluster5.markers <- FindMarkers(EW.2, ident.1 = 5, min.pct = 0.25)
write.csv(cluster5.markers,"./EW2-out/cluster5.markers.csv", row.names = TRUE)
head(cluster5.markers, n = 6)
```

```{r}
cluster6.markers <- FindMarkers(EW.2, ident.1 = 6, min.pct = 0.25)
write.csv(cluster6.markers,"./EW2-out/cluster6.markers.csv", row.names = TRUE)
head(cluster6.markers, n = 6)
```

```{r}
cluster7.markers <- FindMarkers(EW.2, ident.1 = 7, min.pct = 0.25)
write.csv(cluster7.markers,"./EW2-out/cluster7.markers.csv", row.names = TRUE)
head(cluster7.markers, n = 6)
```

```{r}
cluster8.markers <- FindMarkers(EW.2, ident.1 = 8, min.pct = 0.25)
write.csv(cluster8.markers,"./EW2-out/cluster8.markers.csv", row.names = TRUE)
head(cluster8.markers, n = 6)
```

```{r}
cluster9.markers <- FindMarkers(EW.2, ident.1 = 9, min.pct = 0.25)
write.csv(cluster9.markers,"./EW2-out/cluster9.markers.csv", row.names = TRUE)
head(cluster9.markers, n = 6)
```

```{r}
cluster10.markers <- FindMarkers(EW.2, ident.1 = 10, min.pct = 0.25)
write.csv(cluster10.markers,"./EW2-out/cluster10.markers.csv", row.names = TRUE)
head(cluster10.markers, n = 6)
```

```{r}
cluster11.markers <- FindMarkers(EW.2, ident.1 = 11, min.pct = 0.25)
write.csv(cluster11.markers,"./EW2-out/cluster11.markers.csv", row.names = TRUE)
head(cluster11.markers, n = 6)
```

```{r}
cluster12.markers <- FindMarkers(EW.2, ident.1 = 12, min.pct = 0.25)
write.csv(cluster12.markers,"./EW2-out/cluster12.markers.csv", row.names = TRUE)
head(cluster12.markers, n = 6)
```

```{r}
FeaturePlot(EW.2, features = c("MS4A1", "CD19", "CD79A", "TBX21", "CD27", "ITGAX", "CD38", "JCHAIN", "IGHD"))
```

```{r}
pdf("./EW2-out/expression.pdf", width = 9, height = 6)
FeaturePlot(EW.2, features = c("MS4A1", "CD19", "CD79A", "TBX21", "CD27", "ITGAX", "CD38", "JCHAIN", "IGHD"))
dev.off()
```

```{r}
FeaturePlot(EW.2, features = c("IGHG1", "IGHG2", "IGHD", "IGHV4-31", "IGHV4-34", "CD74", "CD83", "IGKV1D-39"))
```

```{r}
EW.2.markers <- FindAllMarkers(EW.2, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
EW.2.markers %>%
    group_by(cluster) %>%
    slice_max(n = 5, order_by = avg_log2FC)
```

```{r}
EW.2.allmarkers <- FindAllMarkers(EW.2, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.25)
EW.2.allmarkers %>%
    group_by(cluster) %>%
    slice_max(n = 5, order_by = avg_log2FC)
```



```{r}
DotPlot(EW.2, features = c("MS4A1", "CD27", "CXCR5", "ITGAX", "CD69", "CD83", "IRF4", "IGHG1", "IGHD", "IGHM", "IGHA1")) + RotatedAxis()
```

## Removing Doublets

```{r}
library(DoubletFinder)
## pK Identification (no ground-truth) 
sweep.res.list <- paramSweep_v3(EW.2, PCs = 1:11, sct = FALSE)
sweep.stats <- summarizeSweep(sweep.res.list, GT = FALSE)
bcmvn <- find.pK(sweep.stats)
```


```{r}
## Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
annotations <- EW.2@meta.data$seurat_clusters
homotypic.prop <- modelHomotypic(annotations)           ## ex: annotations <- seu_kidney@meta.data$ClusteringResults
nExp_poi <- round(0.096*nrow(EW.2@meta.data))  ## doublet formation rate - t10X says 0.8% per 1000 cells - ~12000 cells so 9.6%
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
```


```{r}
## Run DoubletFinder with varying classification stringencies ----------------------------------------------------------------
EW.2 <- doubletFinder_v3(EW.2, PCs = 1:10, pN = 0.25, pK = 0.09, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)
```

EW.2@meta.data[["DF.classifications_0.25_0.09_1137"]]

```{r}
c <- DimPlot(EW.2, pt.size = 1, label=FALSE, label.size = 5, reduction = "umap", group.by = "DF.classifications_0.25_0.09_1137" )

c
```

```{r}
pdf("EW2-out/EW2_doublets.pdf", width = 7, height = 5)
c
dev.off()
```


```{r}
VlnPlot(EW.2, features = c("nCount_RNA", "nFeature_RNA"), group.by = "DF.classifications_0.25_0.09_1137")
```

```{r}
VlnPlot(EW.2, features = c("MKI67", "DNTT"), group.by = "DF.classifications_0.25_0.09_1137")
```


```{r}
EW.2_clean <- subset(EW.2, cells=rownames(EW.2@meta.data)[which(EW.2@meta.data$DF.classifications_0.25_0.09_1137 == "Singlet")])

EW.2_clean
```


```{r}
VlnPlot(EW.2_clean,features="nCount_RNA")
```

```{r}
VlnPlot(EW.2_clean,features="nFeature_RNA")
```



```{r}
DimPlot(EW.2_clean, reduction = "umap")
```


```{r}
table(Idents(EW.2))
```

```{r}
table(Idents(EW.2_clean))
```


## 15. Save Seurat object for downstream analysis

```{r}
saveRDS(EW.2_clean, file = "./EW1-out/EW.2_clean.rds")
```



```{r}
sessionInfo()
```


Convert to anndata
```{r}
library(Seurat)
library(SeuratDisk)

SaveH5Seurat(EW.2_clean, filename = "./EW2-out/EW2.h5Seurat")
Convert("./EW2-out/EW2.h5Seurat", dest = "h5ad")
```